### 2018-04-04 paper.Rmd
finished draft of teaching section

## Child-language class tutorials
## Data-science tutorials

>>>>

## In class demonstrations

Teachers of courses on early language acquisition often want to demonstrate interesting changes in children’s early language. One teaching method is to present static displays on slides or handouts that show (a) text from parent-child conversations extracted from CHILDES or (b) data visualizations of changes in various metrics related to language production and input (e.g., MLU or Frequency). The challenge with static displays or handouts is that they cannot be modified during a lecture and thus rely on the instructor selecting examples that they think will be compelling to their students.

Interactive, in-class demonstrations offer a solution and can be a powerful way to explain complex concepts while increasing student engagement with the course materials. One goal in creating the browser-based interactive visualization tools in childes-db is to provide instructors with a new tool for illustrating key concepts in children's language acquisition. In this section, we walk through one example using childes-db to show the developmental trajectories of different kinds of words that in children's early production. We also highlight ways that instructors could make the in-class demonstration interactive by eliciting hypotheses from students, which become operationalized as queries to the interactive visualization tools.

### Worked example

What kinds of words do children first produce? This question is a good candidate for an interesting "hook" that gets students thinking about some key issues in child language acquisition. Diary studies and large-scale studies using parent report show that children’s first words tend to fall into a fairly small number of categories: people, food, body parts, clothing, animals, vehicles, toys, household objects, routines, and activities or states [@clark2009first; @fenson1994variability]. The key insight is that young children talk about what is going on around them: people they see every day, e.g., toys and small household objects they can manipulate or food they can control.

A teacher could illustrate children's early vocabularies by showing a table with columns for the lexical category, examples of words from that category, and children's average age of production. To make this an interactive demonstration using childes-db, the teacher could:

1. introduce the question: What kinds of words do children first produce?
2. allow students to reflect or do a pair-and-share discussion with their neighbor
3. show the trajectory of a single lexical item while explaining the key parts of the visualization (see Panel A of Fig. XX)
4. elicit hypotheses from students about the kinds of words that children are likely to produce 
5. make real-time queries to the interactive visualization tool to add students' suggestions and talk through the updated plots 
6. finish by entering a pre-selected set of words that communicate the critical takeaway (e.g., see Panel C of Fig. XX)

```{r in-class-demo-flow, fig.pos="h", out.width = "100%", fig.cap="Worked example of using the interactive visualization tools for in-class teaching. Panels A-D show how an instructor could dynamically build a plot during a lecture to interactively demonstrate a key concept in language acquisition."}
library(here)
library(png)
png::readPNG(here::here("imgs/in_class_teach_viz.png")) %>% grid::grid.raster()
```

\noindent
Another benefit of this workflow is that the instructor could easily reuse it to demonstrate other concepts or questions that come up during the lecture, e.g., order of production for different types of words such as quantifiers vs. number words or function vs. content words. Moreover, future work on childes-db interactive visualization tools could add part of speech tags, phonological information, and n-grams. This would allow instructors to use childes-db to demonstrate concepts such as when children start producing certain phonological patterns (e.g., consonant clusters), the debate about the extent of the noun/verb bias in children's early vocabulary, and the development of particular syntactic constructions.

In addition to the interactive data visualizations, instructors can use childes-db to download the raw data used to generate the figures. This allows teachers an opportunity to have students further explore and analyze the data they see in the plots. Moreover, this functionality provides an example of another way childes-db can facilitate language acquisition courses by making data easily-accessible for tutorials and student research projects. In the next section, we walk through one example of how an instructor could use childes-db to create a tutorial that scaffolds students' ability to ask interesting computational questions using child language data.

## Tutorials and assignments

One of the goals for courses on applied natural language processing (NLP) is for students to get hands-on experience using state of the art NLP tools to analyze real-world language data [@hearst2005teaching]. A primary challenge for the instructor is to decide how much time should be spent teaching the requisite programming skills for accessing and formatting language data, which are typically unstructured. One pedagogical strategy is to abstract away these details and avoid having students deal with obtaining data and formatting text. This approach shifts students' effort away from data cleaning and towards programming analyses that encourage students to explore language data and test interesting hypotheses.

In the introduction, we outlined the challenge of getting students started with analyzing data from CHILDES. These barriers include learning the details of a special file format (CHAT) and a command-line tool (CLAN) built to specifically for the CHAT format. This workflow can be daunting to navigate for a beginner. Thus, an instructor would have to spend course time teaching students to use CHAT and CLAN, leaving less time for students to try analyses that might increase their interest in child language acquisition. Moreover,  while CLAN provides a useful tool for interactive exploration,  it does not connect to a pipeline of modern natural language processing tools (e.g., models of word embeddings @mikolov2013distributed), making it difficult for students to explore the data using new NLP tools.

We hope that childes-db combined with the `childesr` API will provide instructors with a solution to both of these challenges. The API gives students an easy-to-learn method for programmatic access to CHILDES data. This will reduce the time spent getting data out of CHILDES and allow more time for thinking about interesting questions that can be answered with NLP and child language data. Second, by writing our API in the R programming language, students will gain access to a general workflow can be integrated into a pipeline that includes the newest NLP tools.

For example, an instructor could create a programming assignment with the goal of reproducing the key findings in the case studies that we presented earlier -- color words or gender. Depending on the students' knowledge of R, the instructor could decide how much of the `childesr` code to provide before asking students to generate their own plots and writeups.

Moreover, the instructor could take advantage of recent advances in tools for teaching literate programming such as R Markdown [@xie2015dynamic; @baumer2014r]. Writing the assignment in an R Markdown notebook would allow the interleaving of prose with executable chunks of code. Instructors could then write prose that is directly related to the course content and the assignment while building in pre-written blocks of executable code that demonstrate key programming concepts for students. To push students' understanding, instructors could write incomplete or blank code blocks that students can use to write their own code. In short, `childesr` combined with R Markdown provides a powerful teaching tool, allowing students to start analyzing language data quickly and providing instructors with a way to build customized lessons for students at different levels of programming experience.
